---
title: "Test Sobol with correlation structure"
author: "Maureen Kennedy"
date: "May 17, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is to test whether the correlation structure created by the Iman method is conserved with a Sobol rearrangement, for global SA.

```{r getData,echo=FALSE,include=FALSE}
# first load in the fit distributions workspace
load("../DistFittingRProj/Workspaces/HurdleCustomFitsNoZeroWNoOut.RData")
set.seed(NULL) # reset the seed, otherwise get same seed every time based on loading this workspace
# read in the mapping between the EVT groups and a characteristic fuel bed
evtFB.map<-read.csv("EVT_Fuelbed_xwalk.csv") # FCCSID is the fuelbed number, EVT_GP is the EVT group
# read in the FCCS fuelbeds
# skip the header line, this is the FCCS fuelbeds file to give baseline values for the fuel loadings
all.fbs<-read.table("consume/input_data/fccs_loadings.csv",skip=1,header=TRUE,sep=",",
                    stringsAsFactors=TRUE) 
#  fueltype names in FCCS with those in database names
fbLoadNames.df<-read.table("FCCS_DataBaseCategoryMapUpdatedWFOFEM.csv",
                           stringsAsFactors = FALSE,sep=",",header=TRUE)

# should contain everything necessary to estimate correlation matrices
# Make sure that the woody totals are present
data.file$X10Khr_loading_Mgha<-data.file$X10KhrR_loading_Mgha+data.file$X10KhrS_loading_Mgha
data.file$X1Khr_loading_Mgha<-data.file$X1KhrR_loading_Mgha+data.file$X1KhrS_loading_Mgha

data.file$cwd_loading_Mgha<-data.file$X10Khr_loading_Mgha+data.file$X1Khr_loading_Mgha+data.file$GT10KhrR_loading_Mgha+data.file$GT10KhrS_loading_Mgha


data.file$fwd_loading_Mgha<-data.file$X1hr_loading_Mgha+data.file$X10hr_loading_Mgha+data.file$X100hr_loading_Mgha

# so now we need a new function, to calculate a complete and proper
# correlation matrix on the subsets of interest in the database
smolder1.id<-c(4,5,8,9,16,17)
smolder2.id<-c(3,7,16,17) # because we don't seem to have a total column for >10K

flaming1.id<-c(1,2,6,18,20,22,23,26)
flaming2.id<-c(15,18,20,22,23,26) # assuming fwd is 1-hr+10-hr+100-hr
flaming3.id<-c(1,2,6,18,22,26) # remove lichen and moss because those data seem to be rare

# now we will subset by the groups for an example EVT, and calculate the
# correlation matrix. It does seem reasonable that if 1 in this group is measured,
# the others likely also are
# try with the first EVT
#replace commented section below with function call
#######
cur.evt<-58 # 
tmp.df<-data.file[data.file[,EVTCol]==evt.vals[cur.evt],]
tmp.loads.df<-tmp.df[,start.col:ncol(tmp.df)]
# might need an imputation package to assess the missingness 
# in the subset of data
tmp2.loads.df<-tmp.loads.df[,smolder1.id]

# package mice has a way to assess the missingness in a matrix
library(mice)
tmp.md<-md.pattern(tmp2.loads.df)

N.samp<-5000

source("../Functions/SimHurdleDist_FN.R")
source("../Functions/SensitivitySampleWithCorr_FN.R")
sens.mats.smolder.list<-corr.sa.fn(data.file,fuel.ids=smolder1.id,complete.case=TRUE,
                                   min.co.occur=30,
                     evts=evt.vals,EVTCol=EVTCol,start.col=start.col,n.samp=N.samp,
                     rankObj=distributionCustomRankingHurdleNOut,
                     fitObj=distributionCustomFittingHurdleNOut)

sens.mats.smolder.list.upper<-corr.sa.fn(data.file,fuel.ids=smolder1.id,complete.case=TRUE,
                                         min.co.occur=30,
                     evts=evt.vals,EVTCol=EVTCol,start.col=start.col,n.samp=N.samp,
                     rankObj=distributionCustomRankingHurdleNOut,
                     fitObj=distributionCustomFittingHurdleNOut,
                     upper.quantile = 0.95)
# looks like lichen and moss might present an issue here. Try without
# might also be more of an issue with complete cases
sens.mats.flame.list<-corr.sa.fn(data.file,fuel.ids=flaming3.id,complete.case=TRUE,
                                 min.co.occur=30,
                     evts=evt.vals,EVTCol=EVTCol,start.col=start.col,n.samp=N.samp,
                     rankObj=distributionCustomRankingHurdleNOut,
                     fitObj=distributionCustomFittingHurdleNOut)
sens.mats.flame.list.upper<-corr.sa.fn(data.file,fuel.ids=flaming3.id,complete.case=TRUE,
                                       min.co.occur=30,
                     evts=evt.vals,EVTCol=EVTCol,start.col=start.col,n.samp=N.samp,
                     rankObj=distributionCustomRankingHurdleNOut,
                     fitObj=distributionCustomFittingHurdleNOut,
                     upper.quantile=0.95)

# this yields a list, where the first component is the sampled values in a dataframe
# if you set upper.quantile to a value, then the sample is truncated to the upper quantile
library(sensitivity)
sobolF<-sobolEff(X1=sens.mats.flame.list[[1]][[cur.evt]][1:(N.samp/2),],X2=sens.mats.flame.list[[1]][[cur.evt]][(N.samp/2+1):N.samp,],nboot = 1000)
cor(sobolF$X)
cor(sens.mats.flame.list[[1]][[cur.evt]])
# seems to keep the general form of the correlation matrix
# so now we can run the SA as before--keeping everything else at the baseline values for the given fuelbed
# change base EVT for smoldering distribution coverage
# note that these will have a lot of zeroes, so
# might need to change sampling scheme
#sobolS<-sobolEff(X1=sens.mats.smolder.list[[1]][[cur.evt]][1:500,],
#                 X2=sens.mats.smolder.list[[1]][[cur.evt]][501:1000,],
#                 nboot = 1000)
```

First generate the correlated sample, and then create the sobol object. Identify baseline values for all other necessary fuel types, and add as constant to final dataframe. Call consume (or fofem), then evaluate results. we will do this with an example evt

```{r makeConsumeInputFromSobol, echo=FALSE,include=FALSE}
source("GenerateConsumeLoadingsFileCorrSamp.R")
cur.evt.num<-evt.vals[cur.evt]
  base.fb<-evtFB.map$FCCSID[evtFB.map$EVT_GP==cur.evt.num]#
  base.fb<-unique(base.fb[!is.na(base.fb)])#&!is.na(evtFB.map$FCCSID)][1]

  corr.sampF.vals<-data.frame(sobolF$X)
  names(corr.sampF.vals)<-names(sens.mats.flame.list[[1]][[cur.evt]])
  consume.fuel.loads<-GenerateSobolFuelInput.fn(corr.sampF.vals,nreps,
                                                fbLoadNames.df,all.fbs,base.fb,change.units=T)
# for prcc  
  consume.fuel.loads<-GenerateFuelInput.fn(sens.mats.flame.list[[1]][[cur.evt]],nreps,
                                                fbLoadNames.df,all.fbs,base.fb,change.units=T)
  outfilename<-"FuelLoadInputSA.csv"
  loadInFile.head<-matrix(c("GeneratorName=FCCS   3.0","GeneratorVersion=3.0.0","DateCreated=07/18/2016"),ncol=3)
  # didn't like today's date
      # write the file with the proper header information
  write.table(loadInFile.head,file=outfilename,row.names=FALSE,col.names=FALSE,sep=",",quote = FALSE)
      # Now append that file with the generated loadings
  write.table(consume.fuel.loads,file=outfilename,row.names=FALSE,append=TRUE,sep=",",quote=FALSE)

      # and create an environmental input file that is just repeated for all rows of the fuels file
      
  env.in<-read.csv("sample_consume_input.csv") ###Note, if this file changes so should the FOFEM input file moistures
  new.env.in<-env.in[1,]
  for(m in 2:nrow(consume.fuel.loads))
  {
    new.env.in<-rbind(new.env.in,env.in[1,])
  }
  new.env.in$fuelbeds<-consume.fuel.loads$fuelbed_number
  envoutfilename<-"EnvInputSA.csv"
  write.table(new.env.in,file=envoutfilename,row.names=FALSE,sep=",")

```

Next we call the consume batch processor. We can do this from R.

```{r callConsume, echo=FALSE, include=FALSE}
  system.call<-paste("python consume_batch.py natural",envoutfilename,  "-f", outfilename)
  try1<-try(system(system.call)) # tells R to execute this system call in the working directory
      
  if(try1==1)
  {
    print("Failed consume call")
  }
  if(try1==0)
  {
    results.sa<-read.csv("consume_results.csv") # writes to this file every time, replacing previous    
  }
results

```

And now we calculate the sensitivity metrics

```{r CalcSens, echo=FALSE,fig.height=12,fig.width=8}
response.vars<-c("PM.Emissions","CO.Emissions","CO2.Emissions","PM25.Emissions")
par(mfrow=c(2,2),mar=c(12,3,1.5,0.5),mgp=c(2,0.5,0))
sens.list<-list()
for(k in 1:length(response.vars))
{ # here we standardize outputs to the z-scoore
  sens.list[[k]]<-tell(sobolF,
                       (results.sa[,response.vars[k]]-mean(results.sa[,response.vars[k]]))/sd(results.sa[,response.vars[k]]))
  rownames(sens.list[[k]]$S)<-names(corr.sampF.vals)
  plot(1:nrow(sens.list[[k]]$S),sens.list[[k]]$S$original,axes=FALSE,ylim=c(-.15,1),pch=16,
             xlab="",ylab="Sensitivity index",
       main=paste("EVT:",cur.evt,"FB",base.fb,"output:",response.vars[k]))#las=2)
        
        axis(1,at=1:nrow(sens.list[[k]]$S),labels=rownames(sens.list[[k]]$S),
             las=2,cex.lab=0.9,tick = F)
        segments(x0 = 1:nrow(sens.list[[k]]$S),x1=1:nrow(sens.list[[k]]$S),
                 y0=sens.list[[k]]$S$`min. c.i.`,y1=sens.list[[k]]$S$`max. c.i.`)
        axis(2,las=1)
        abline(h=0,lwd=2,col="grey") # to see if CI for SI is above 0
        box()
}

``` 

Now let's try to create a FOFEM input file from the Sobol sample, with the switch indicated large fuel loads by size class.

```{r fofemSobol, echo=FALSE,include=FALSE}
#Assumes that the sobol sample was already generated for consume
# this file fill guide the creation of the FOFEM input file
# we will need to update with the appropriate moistures
# regions and emissions depending on the base EVT for the SA
setwd("fofem")
base.fofem<-read.csv("FOFEM_6_May12013_MCKUpdate2.csv") 
# note, this file has the header--our final file 
# note also this has the fuel moistures matched to consume env input as
# best we can. If that changes, so should this
base.fofem.use<-base.fofem[1,]

  fofem.fuel.loads<-GenerateSobolFuelInput.fn(corr.sampF.vals,nreps,fbLoadNames.df,
                                                all.fbs,base.fb,base.fofem=base.fofem.use,change.units=T,mod="F")
  
# write it to an input file for FOFEM
# First write the switch for expanded CWD
fofem.filename<-"FOFEM_FlamingSAInput1.csv"
write("#1k-SizeClass",file=fofem.filename)
write.table(fofem.fuel.loads,file=fofem.filename,append = TRUE,sep=",",row.names = FALSE,col.names = FALSE)
system.call<-paste("FOF_GUI C",fofem.filename,"ConE-Out.txt ConE-run.txt ConE-Err.txt H", sep=" ")
system(system.call) # tells R to execute this system call in the working directory

```

Note: we're getting a warning that no ignition simulated for some of the stands. Need to look into that. I think it was because of the relatively high fuel moistures. Goes away when the fm's match the consume environmental inputs. 

```{r plotFOFEMSA,echo=FALSE,fig.height=12,fig.width=8}
check.fofem<-scan("fofem/ConE-Err.txt",what = character())# if there is no error recorded, then this file will be empty

if(length(check.fofem)==0) # if there is no error recorded, then this file will be empty
{
  results.fofem.sa<-read.csv("fofem/ConE-Out.txt") # writes to this file every time, replacing previous results

  # for flaming:
  responseF.vars<-c("COF","CO2F","PM25F")

  # for smoldering:
  responseS.vars<-c("COS","CO2S","PM25S")
  par(mfrow=c(2,2),mar=c(12,3,1.5,0.5),mgp=c(2,0.5,0))
  sens.list<-list()
  for(k in 1:length(responseF.vars))
  { # here we standardize outputs to the z-scoore
    sens.list[[k]]<-tell(sobolF,
                         (results.fofem.sa[,responseF.vars[k]]-mean(results.fofem.sa[,responseF.vars[k]]))/sd(results.fofem.sa[,responseF.vars[k]]))
    rownames(sens.list[[k]]$S)<-names(corr.sampF.vals)
    plot(1:nrow(sens.list[[k]]$S),sens.list[[k]]$S$original,axes=FALSE,ylim=c(-.15,1),pch=16,
               xlab="",ylab="Sensitivity index",
         main=paste("EVT:",cur.evt,"FOFEM","output:",responseF.vars[k]))#las=2)
          
          axis(1,at=1:nrow(sens.list[[k]]$S),labels=rownames(sens.list[[k]]$S),
               las=2,cex.lab=0.9,tick = F)
          segments(x0 = 1:nrow(sens.list[[k]]$S),x1=1:nrow(sens.list[[k]]$S),
                   y0=sens.list[[k]]$S$`min. c.i.`,y1=sens.list[[k]]$S$`max. c.i.`)
          axis(2,las=1)
          abline(h=0,lwd=2,col="grey") # to see if CI for SI is above 0
          box()
  }

}
# else
# {
#   print("Check FOFEM error file!")
#   
# }
      

  
```

air quality standard for e.g., PM 2.5 is 15 micrograms/m^3. We are given lbs/acre (at least for fofem, probably consume as well)

1 lb = 453600000 micrograms

1 acre foot = 1233.48 cubic meters

lbs/acre foot * 453600000 micrograms/lb * 1 acre foot / 1233.48 m^3